---
title: "covers scrape"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load packages
library(tidyverse)
library(rvest)
library(shiny)
library(rsconnect)
```

```{r}
#read csv to get school names
covers <- read_csv("data/covers_scrape_ids.csv")
#6 teams missed due to rpi change
missed <- tibble(Team = c("duke-blue-devils", "tennessee-volunteers", "stephen-f.-austin-lumberjacks", "vcu-rams", "wright-st.-raiders"),
                 Season = c("2015-2016", "2008-2009", "2007-2008", "2016-2017", "2009-2010"))

covers <- convers %>% bind_rows(missed)

#need to do separate scrape for after 2015-2016 (number of tables on covers changes)
scrape1 <- c("2006-2007","2007-2008","2008-2009","2009-2010","2010-2011","2011-2012","2012-2013","2013-2014", "2014-2015", "2015-2016")
covers <- covers %>% filter(Season %in% scrape1)

scrapeStatTables <- function(school,season){
  url <- paste0("https://www.covers.com/sport/basketball/ncaab/teams/main/",school,"/",season,"/")
  print(url)
  webpage <- read_html(url)
  webpage %>% html_table(fill = TRUE) -> tables
  return(tables)
}
```


```{r}
#create tibbles for betting data
reg_season_betting_data <- tibble()
postseason_betting_data <- tibble()

#runs scrape in for loop, for each school name and season, scrape that school's schedule w/ betting data, fill tibble with tables, add school name
for(i in 1:626){
    Sys.sleep(runif(1)*7)
    teamTables <- tail(scrapeStatTables(covers[i,]$team, covers[i,]$Season),3)
    
    regular_season <- teamTables[[2]]
    regular_season$Team <- covers[i,]$team
    regular_season$Season <- covers[i,]$Season
    
    postseason <- teamTables[[1]]
    postseason$Team <- covers[i,]$team
    postseason$Season <- covers[i,]$Season
    
    postseason_betting_data %>% bind_rows(postseason) -> postseason_betting_data
    reg_season_betting_data %>% bind_rows(regular_season) -> reg_season_betting_data
}
```

#find teams from scrape1 that collected player data instead
```{r}
broken <- reg_season_betting_data[is.na(reg_season_betting_data$`Regular Season...1`),] %>% group_by(Team, Season) %>% summarise(n = n()) 

broken <- broken %>% select(c(1:2))
names(broken)[1] <- "team"
```

#add broken teams from scrape 1 to post 2015-2016 teams
```{r}
covers <- read_csv("data/covers_scrape_ids.csv")
scrape2 <- c("2016-2017","2017-2018","2018-2019")
covers <- covers %>% filter(Season %in% scrape2)
covers <- covers %>% bind_rows(broken)
```


#scrape 2 for post 2015-2016 and broken teams from scrape 1
```{r}
#create tibbles for betting data
reg_season_betting_data_post16 <- tibble()
postseason_betting_data_post16 <- tibble()

#runs scrape in for loop, for each school name and season, scrape that school's schedule w/ betting data, fill tibble with tables, add school name
for(i in 1:626){
    Sys.sleep(runif(1)*7)
    teamTables <- tail(scrapeStatTables(covers[i,]$team, covers[i,]$Season),5)
    
    regular_season <- teamTables[[2]]
    regular_season$Team <- covers[i,]$team
    regular_season$Season <- covers[i,]$Season
    
    postseason <- teamTables[[1]]
    postseason$Team <- covers[i,]$team
    postseason$Season <- covers[i,]$Season
    
    postseason_betting_data_post16 %>% bind_rows(postseason) -> postseason_betting_data_post16
    reg_season_betting_data_post16 %>% bind_rows(regular_season) -> reg_season_betting_data_post16
}
```

#find broken teams from scrape 2
```{r}
#uconn had postseason ban in 2013, must remove them from sample
broken2 <- reg_season_betting_data_post16[is.na(reg_season_betting_data_post16$`Regular Season...1`),] %>% group_by(Team, Season) %>% filter(Team != "connecticut-huskies") %>% summarise(n = n())
```

#scrape broken teams from scrape 2 (scrape 3)
```{r}
#create tibbles for betting data
reg_season_betting_data_broken2 <- tibble()
postseason_betting_data_broken2 <- tibble()

#runs scrape in for loop, for each school name and season, scrape that school's schedule w/ betting data, fill tibble with tables, add school name
for(i in 1:626){
    Sys.sleep(runif(1)*7)
    teamTables <- tail(scrapeStatTables(broken2[i,]$Team, broken2[i,]$Season),3)
    
    regular_season <- teamTables[[2]]
    regular_season$Team <- broken2[i,]$Team
    regular_season$Season <- broken2[i,]$Season
    
    postseason <- teamTables[[1]]
    postseason$Team <- broken2[i,]$Team
    postseason$Season <- broken2[i,]$Season
    
    postseason_betting_data_broken2 %>% bind_rows(postseason) -> postseason_betting_data_broken2
    reg_season_betting_data_broken2 %>% bind_rows(regular_season) -> reg_season_betting_data_broken2
}
```

#clean data from scrap 1 and scrape 2 to remove broken teams, combine all 3 scrapes into one table for reg season and one table for postseason
```{r}
#regular season
reg_betting_data <- reg_season_betting_data %>% filter(!is.na(`Regular Season...1`)) %>% select(c(1:5,8,9))

scrape2_data <- reg_season_betting_data_post16 %>% filter(!is.na(`Regular Season...1`)) %>% select(c(1:5,8,9))

scrape3_data <- reg_season_betting_data_broken2 %>% select(c(1:5,8,9))

reg_betting_data <- reg_betting_data %>% bind_rows(scrape2_data, scrape3_data)

#postseason
post_betting_data <- postseason_betting_data %>% filter(!is.na(Playoffs...1)) %>% select(c(1:5,9,10))

post_scrape2_data <- postseason_betting_data_post16 %>% filter(!is.na(Playoffs...1)) %>% select(c(1:5,9,10))

post_scrape3_data <- postseason_betting_data_broken2 %>% select(c(1:5,9,10))

post_betting_data <- post_betting_data %>% bind_rows(post_scrape2_data, post_scrape3_data)
```

#change column names in both tables and combine into one table
```{r}
names(reg_betting_data)[1] <- "date"
names(reg_betting_data)[2] <- "opponent"
names(reg_betting_data)[3] <- "score"
names(reg_betting_data)[4] <- "ats"
names(reg_betting_data)[5] <- "over_under"

names(post_betting_data)[1] <- "date"
names(post_betting_data)[2] <- "opponent"
names(post_betting_data)[3] <- "score"
names(post_betting_data)[4] <- "ats"
names(post_betting_data)[5] <- "over_under"

betting_data <- bind_rows(reg_betting_data, post_betting_data)

#remove header rows from data and order columns
betting_data <- betting_data %>% filter(date != "Date")
betting_data <- betting_data %>% select(c(Team, Season, date, opponent, score, ats, over_under))
```

#separate/clean columns, export to csv
```{r}
betting_data_clean <- betting_data %>% separate(ats, into = c("result_ats", "spread"), sep = " ")
betting_data_clean2 <- betting_data_clean %>% separate(date, into = c("month", "day"), sep = " ")

betting_data_clean2 %>% write_csv("data/betting_data.csv")
```
